---
title: "Introduction to tidymar"
author: "Hersh Gupta"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE,message=FALSE,warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  message = FALSE,
  warning = FALSE
)
library(dplyr)
library(tidyr)
library(purrr)
library(tidymar)
```

The Office of the Chief Technology Officer for DC provides the [DC GIS Master Address Repository](https://dcatlas.dcgis.dc.gov/mar/) (MAR), available for the public to use. The MAR consists of various applications that rely on web services to find and verify DC-based locations. DC agencies may use the MAR Geocoder application to find, clean, and verify addresses. 

Researchers, programmers, and data scientists may be more inclined to use [MAR web services](https://opendata.dc.gov/pages/mar-webservices) directly to reduce overhead from the Geocoder application and ensure reproducibility in analyses. Users can [query MAR web services using common HTTP methods](https://opendata.dc.gov/pages/mar-webservices#ember104). 

The `tidymar` R package makes this process easier by providing pre-built functions corresponding to the services available through the MAR web services. This document introduces the `tidymar` package.

## Basic Usage 

To get started with the `tidymar` package, users should load it first.

```{r loading, eval=T,echo=T}
library(tidymar)
```

The main function currently implemented in `tidymar` is `find_location`. 

This function serves as a catch-all for all verified DC locations, as it not only geocodes the address, but provider additional information about the location, such as Police Service Area (`PSA`), Neighborhood Cluster (`CLUSTER_`), Voter Precinct (`VOTE_PRCNT`), whether it has condo units or residential units (`HAS_CONDO`, `HAS_RES_UNIT`), and the confidence level (`CONFIDENCE_LEVEL`) of the matched location.

In this basic example, we get the full address and coordinates of the White House using simply "white house" as the input string. The results are in a wide tibble format, which I convert to long and print the full results below:

```{r example1, eval=T,echo=T}
find_location("white house") %>% gather() %>% print(n = Inf)
```

The full MAR Data Dictionary is [available here](https://octo.dc.gov/sites/default/files/dc/sites/octo/publication/attachments/DCGIS_MarDataDictionary_0.pdf).

## Batch Geocoding

Let's say we have a list of DC addresses, (`locations`) and would like to geocode them. 

```{r example2a, eval=T,echo=T}
locations <- tibble(addr = c("899 North Capitol St NE", "1350 Pennsylvania Ave NW", "430 E St NW", "1600 Pennsylvania Ave NW", "1 First St NE"), 
                    name = c("DC Department of Health", "John Wilson Building", "District of Columbia Court", "The White House", "US Supreme Court"), 
                    type = c("local","local","local","federal","federal"))

locations
```

The `find_location` can be vectorized over multiple addresses and returns all possible matching location data. It can also accept intersections, streets, and blocks. 

```{r example2b, eval=T,echo=T}
find_location(locations$addr)
```


